Viscous Transport in Eroding Porous Media

### 2019-10-14 docs/transport/erosion_transport.tex
started changes suggested by Nick. Took off two pages.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:intro}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Porous media flow plays an important role in many environmental and
industrial applications.  Depending on the application, length scales
can vary from $10^{-6}\mathrm{m}$ to
$10^{-1}\mathrm{m}$~\citep{mil-chr-imh-mcb-ped1998} and velocity scales
can be as small as $10^{-1}\mathrm{m/day}$~\citep{kut-scr-dav-ham1995}.
Moreover, for a single porous geometry, the pore sizes and velocities
can range over several orders of magnitude.  Numerical methods that
resolve this range of scales offer the ability to: (i) characterize
dispersion~\citep{saf1959}, (ii) quantify
mixing~\citep{leb-den-dav-bol-car-dec-bou2011, den-leb-eng-bij2011}, and
(iii) develop meaningful constitutive relationships that link the
microscopic and macroscopic realms~\citep{mil-chr-imh-mcb-ped1998}.
Examples of coarse-grained models for porous media flow include
pressure-saturation-permeability
relationships~\citep{mil-chr-imh-mcb-ped1998}, permeability-porosity
relationships~\citep{dar-mcc1998, car1937, won-kop-tom1984}, first
passage time distributions~\citep{ber-sch-sil2000, hym-den-hag-kan2019,
cve-che-wen1996}, tortuosities~\citep{hak-com-den2019, mat-kha-koz2008,
dud-koz-mat2011, kop-kat-tim1996}, concentration
distributions~\citep{ica-den2019, bel-sal-rin1992}, pore and flow
distributions~\citep{ali-par-wei-bre2017}, geometry
connectivity~\citep{knu-car2005, wes-blo-gra2001}, anomalous
dispersion~\citep{dea-qua-bir-jua2018, den-cor-sch-ber2004,
sie-ili-pri-riv-gua2019}, and more.

\todo[inline]{The above set of references for ``examples of course-grained models" seems very thorough and perhaps not as relevant as other references on erosion, melting, etc. JFM's reference format makes them take up a lot of space. So if there are references here that you wouldn't mind dropping, it might help. I wouldn't drop all of them. I think at least 3-4 are definitely needed.}

@@

Owing to the scales present in groundwater flow~\citep{bea1972}, we
model the hydrodynamics with the two-dimensional incompressible Stokes
equations.  Meanwhile, individual grains erode at a rate proportional to
the hydrodynamic shear stress~\citep{wan-fel2004,
ris-moo-chi-she-zha2012, moore2013self, par-izu2000}.  We compute the
vorticity in the fluid bulk since, on solid boundaries, vorticity
reduces to shear and thus provides a convenient way to simultaneously
visualize local erosion rates and changes in the surrounding flow
(figure~\ref{fig:Eroding50vort}).  A variety of numerical methods for
the fluid solver are available including lattice-gas or lattice
Boltzmann~\citep{kop-kat-tim1996, dar-mcc1998}, finite
volumes~\citep{suo-liu-gan2019, den-ica-hid2018,
sie-ili-pri-riv-gua2019}, finite differences~\citep{leb-ded-dav-bou2007,
knu-car2005}, and cellular automata~\citep{rot1988}.  
\todo[inline]{The set of references above seems to me like an unneeded `straw-man' argument. i.e.~Just because we are using a boundary integral method, does that mean we have to list every other possible way of solving the fluid equations? It takes up a fair amount of space. On the other hand, if there are references here you want to keep, by all means do it.}
However, simulating erosion requires a method that is high-order, can resolve
nearly-touching arbitrarily shaped grains, and can accurately compute
derivatives of the velocity field.  Since the governing equations are
linear and homogeneous, we address these challenges by converting the
fluid equations to a boundary integral equation (BIE).

@@

To compute stable simulations of erosion, we use 
\nick{methods of high order in both space and time.}
The time integration is unchanged from
previous work~\citep{qua-moo2018}.  We apply a mild regularization and a
smoothing term to eliminate numerical instabilities that can be
triggered by changes in sign of the shear stress, and we use a stable
second-order Runge-Kutta method applied to the {\thL}
coordinates~\citep{hou-low-she1994} of the eroding grains.  In this
work, we introduce a new quadrature method to resolve dense suspensions.
The accuracy of the trapezoid rule, which was used in previous work, is
adequate for bodies that are sufficiently separated~\citep{tre-wei2014}, 
but not for grains in close contact.  
One of the earliest quadrature methods for nearly-singular
integrands was developed by~\citet{bak-she1986}, and in recent years,
many other \nick{\st{near-singular integration}} schemes have
followed~\citep{oja-tor2015, kli-tor2018, hel-oja2008a, bea-yin-wil2016,
bea-lai2001, klo-bar-gre-one2013}.  We use a Barycentric quadrature
method~\citep{bar2014, bar-wu-vee2015} since it is a non-intrusive
modification of the trapezoid rule, and the error is guaranteed to be
uniformly bounded.  We extend the original quadrature method to compute
the velocity gradient, which is needed to evaluate the shear stress and
the fluid vorticity.

To characterize transport through the resulting configurations, particle
trajectories must be computed. Depending on the application, microscale
transport can be modelled as pure advection~\citep{dea-qua-bir-jua2018,
leb-ded-dav-bou2007, cve-che-wen1996, puy-gou-den2019},
advection-diffusion~\citep{cus-hu-den1995, dag1987, den-ica-hid2018}, or
with a random walk~\citep{saf1959, bij-blu2006, ber-sch-sil2000}.  In
this work, we assume trajectories to be governed by pure advection
(ie.~no diffusion), so the particle trajectories are identical to the
streamlines. We compute trajectories $\ss(t)$ that are initialized at
$\ss_0$ by solving the advection equation
\begin{align}
  \frac{d\ss}{dt} = \uu(\ss,t), \quad \ss(0) = \ss_0,
  \label{eqn:tracers}
\end{align}
\todo[inline]{Could the above equation be changed to inline to save space?}
where $\uu$ is the fluid velocity.  Since there is no stiff diffusive
term, we solve~\eqref{eqn:tracers} with a fourth-order explicit
Runge-Kutta time stepping method. The Barycentric
quadrature rule is used to accurately compute trajectories that are
close to an eroding grain. 
%We note that an alternative approach to
%compute trajectories is to form a distribution of velocities $\uu$ and
%use a continuous time random walk~\citep{mon-wei1965,
%den-cor-sch-ber2004, leb-den-car2008, ber-cor-den-sch2006}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Governing Equations}
\label{sec:formulation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We start by defining the main variables used to model erosion.  We only
briefly summarize the model, and a more detailed description can be
found in previous work~\citep{qua-moo2018}.  We consider flows inside a
confined geometry $\Omega$ that contains $M$ eroding bodies with
boundaries $\gamma_\ell$, $\ell = 1,\ldots,M$.  The boundary of the
fluid domain is $\bd \Omega = \Gamma \cup \gamma_1 \cup \cdots \cup
\gamma_M$, where $\Gamma$ is the outer boundary, taken to be a slightly
smoothed version of the boundary of $[-3,3] \times [-1,1]$.  We only
place eroding bodies in $[-1,1] \times [-1,1]$ to create a buffer region
that allows the flow profile imposed at the inlet to transition to the more
complex flow intervening between the bodies. On each eroding body, $\nn$
is the normal vector pointing into the body, and $\ss$ is the unit
tangent vector pointing in the counterclockwise direction.  The normal
vector on $\Gamma$ is the outward unit normal.  Neglecting inertial
forces, the governing equations are
\begin{equation}
\label{eqn:erosionModel}
  \begin{split}
    \mu \Delta \uu = \grad p, &\hspace{20pt} \xx \in \Omega, \gap 
      &&\mbox{\em conservation of momentum}, \\
    \grad \cdot \uu = 0, &\hspace{20pt} \xx \in \Omega, \gap 
      &&\mbox{\em conservation of mass}, \\
    \uu = \mathbf{0}, &\hspace{20pt} \xx \in \gamma, \gap 
      &&\mbox{\em no slip on the eroding bodies}, \\
    \uu = \UU, &\hspace{20pt} \xx \in \Gamma, \gap 
      &&\mbox{\em outer wall velocity}, \\
    \Vn = \CE \, \abs{\tau}, &\hspace{20pt} \xx \in \gamma,
      &&\mbox{\em erosion model}.
  \end{split}
\end{equation}
Here, $\uu$ is the fluid velocity, $p$ is the pressure, $\UU$ is a
prescribed Hagen-Poiseuille velocity field, $\Vn$ is the normal velocity
of $\gamma$, and
\begin{align}
  \tau = -(\nabla \uu + \nabla \uu^T) \nn \cdot \ss
  \label{eqn:shearStress}
\end{align}
is the shear stress on $\gamma$. We simulate erosion by alternating
between solving the fluid equations and advancing the eroding grains.
The strength of $\UU$ is adjusted at each time step to achieve a
constant pressure drop across the channel, 
\nick{motivated by the geological situation of a porous medium connecting two regions of  fixed hydraulic heads.}

@@

Once the velocity is computed in $\Omega$, the tortuosity can be
computed with the Eulerian velocity field~\eqref{eqn:tortuosity2}.  We
compute the velocity at $\xx_{ij} = (-1 + i\Delta x, -1 + j\Delta y)$,
$i,j=1,\ldots,N$, where $\Delta x = \Delta y = 2/N$, and the velocity at
points inside an eroding body are assigned a value of 0.  Then, the
tortuosity is approximated as
\begin{align}
  T = \frac{\sum\limits_{i,j=1}^N \|\uu(\xx_{ij})\|\Delta x \Delta y}
      {\sum\limits_{i,j=1}^N u_1(\xx_{ij}) \Delta x \Delta y}.
\end{align}

\todo[inline]{The above vertical fraction could be made into a horizontal fraction, which would save a lot of space.}

@@

Again, given a complex-valued density function $\eta$, the boundary data
of $v[\eta](x)$, as defined in equation~\eqref{eqn:laplaceComplex},
satisfies the Sokhotski-Plemelj jump relation~\eqref{eqn:SPrelation}.
Since the limiting boundary data of $v$ differs when considering $x \in
\Omega$ and $x \in \Omega^c$, we denote the boundary data as $v^-$ for
$x \in \Omega$, and as $v^+$ for $x \in \Omega^c$.  Rather than directly
applying the trapezoid rule to approximate $v(x)$ in
equation~\eqref{eqn:cauchy}, we start with the identity
\begin{align}
  \int_{\bd\Omega} \frac{v^{-}(y) - v(x)}{y - x} \,dy = 0,
    \quad x \in \Omega.
\end{align}
Since the integrand is bounded for all $x \in \Omega$, we can apply the
trapezoid rule
\begin{align}
  \sum_{j=1}^{N} \frac{v^{-}(y_j) - v(x)}{y_j - x} w_j \approx 0,
\end{align}
and the error is independent of $x$.  Rearranging for $v(x)$, we have
the interior Barycentric quadrature rule
\begin{align}
  v(x) = \frac{\sum\limits_{j=1}^N \frac{v^{-}(y_j)}{y_j - x} w_j}
  {\sum\limits_{j=1}^N \frac{1}{y_j - x} w_j}, \quad x \in \Omega.
  \label{eqn:BaryvInterior}
\end{align}
Using a similar construction and letting $a$ be any point inside
$\Omega$, the exterior Barycentric quadrature rule is
\begin{align}
  v(x) = \frac{1}{x-a} 
    \frac{\sum\limits_{j=1}^N \frac{v^+(y_j)}{y_j - x}w_j}
    {\sum\limits_{j=1}^N \frac{(y_j - a)^{-1}}{y_j - x}w_j},
    \quad x \in \Omega^c.
  \label{eqn:BaryvExterior}
\end{align}

\todo[inline]{More prime examples of vertical fractions that could be made horizontal.}

Similar constructions can be used to derive quadrature rules
for $v'(x)$.  For $x \in \Omega$, we have the identity
\begin{align}
  \int_{\bd\Omega} \frac{v^-(y) - v(x) + (y-x)v'(x)}{(y - x)^2} = 0,
\end{align}
and the integrand is bounded for all $x \in \Omega$.  Therefore, after
applying the trapezoid rule and rearranging for $v'(x)$, we have the
interior Barycentric quadrature rule
\begin{align}
  v'(x) = \frac{\sum\limits_{j=1}^{N}
    \frac{v^{-}(y_j) - v(x)}{(y_j-x)^2} w_j}
  {\sum\limits_{j=1}^{N} \frac{1}{y_j-x} w_j}, 
  \quad x \in \Omega.
  \label{eqn:BaryvprimeInterior}
\end{align}
Using a similar construction, the exterior Barycentric quadrature rule
for the the first derivative is
\begin{align}
  v'(x) = \frac{1}{x-a} \frac{\sum\limits_{j=1}^N
    \frac{v^+_j - v(x)}{(y_j - x)^2} w_j}
    {\sum\limits_{j=1}^N \frac{(y_j-a)^{-1}}{y_j - x} w_j},
    \quad x \in \Omega^c.
  \label{eqn:BaryvprimeExterior}
\end{align}
Note that $v(x)$ is required to compute $v'(x)$ for both the interior
and exterior case, and this is available using the Barycentric
quadrature rules~\eqref{eqn:BaryvInterior}
and~\eqref{eqn:BaryvExterior}.

To compute the shear stress and vorticity, we require a Barycentric
quadrature rule for $v''(x)$.  The derivation is largely based on the
work of~\citet[see equation (2.12)]{ioa-pap-per1991}.  We start with
the second derivative of the Cauchy integral theorem
\begin{align}
  0 = \frac{1}{2\upi i} \int_{\bd\Omega} 
      \frac{2v^{-}(y)}{(y-x)^3}\,dy - v''(x).
\end{align}
For the interior case, $x \in \Omega$, we use the identity
\begin{align}
  \frac{1}{2\upi i}\int_{\bd\Omega} \frac{1}{(y-x)^n}\, dy = 
  \left\{
    \begin{array}{ll}
      1, & n = 1, \\
      0, & n = 2,3,\ldots. \\
    \end{array}
  \right.
\end{align}
Combining this identity with the Cauchy integral representation of
$v''(x)$, we have
\begin{align}
  0 = \frac{1}{2\upi i} \int_{\bd\Omega} 
      \frac{2v^{-}(y) - v''(x)(y-x)^2 - 2v(x) - 2(y-x)v'(x)}
      {(y-x)^3}\, dy.
\end{align}
This integrand is constructed so that it is bounded for all $x \in
\Omega$, and applying the trapezoid rule, we have
\begin{align}
  0 \approx  \sum_{j=1}^{N} 
      \frac{2v^{-}(y_j) - v''(x)(y_j-x)^2 - 2v(x) - 2(y_j-x)v'(x)}
      {(y_j-x)^3} w_j,
\end{align}
where the accuracy is independent of $x$.  Solving for $v''(x)$, the
Barycentric quadrature rule for the interior second derivative is
\begin{align}
  v''(x) &\approx \frac{2\sum\limits_{j=1}^N 
    \frac{v^{-}_{j} - v(x) - (y_j-x)v'(x)}{(y_j-x)^3}w_j}
    {\sum\limits_{j=1}^N \frac{1}{y_j-x}w_j}, \quad x \in \Omega.
\end{align}
For the exterior case, $x \in \Omega^c$, we start with the identity
\begin{align}
\frac{1}{x-a} &= -\frac{1}{2\upi i}\int_{\bd\Omega} 
    \frac{(y-a)^{-1}}{y-x}\, dy. 
\end{align}
Combining this identity with the Cauchy integral representation of
$v''(x)$, we have
\begin{align}
  0 = \frac{1}{2\upi i} \int_{\bd\Omega} 
    \frac{2v^+(y) - 2v(x) - 2(y-x)v'(x) + (y-a)^{-1} (x-a) (y-x)^2 v''(x)}
    {(y-x)^3}.
\end{align}
As in the interior case, the integrand is chosen so that it is bounded
for all $x \in \Omega^c$.  Therefore, after applying the trapezoid rule
and solving for $v''(x)$, we have the Barycentric quadrature rule for
the exterior second derivative
\begin{align}
  v''(x) &\approx \frac{1}{x-a}\frac{2\sum\limits_{j=1}^N
    \frac{v^{+}_{j} - v(x) - (y_j-x)v'(x)}{(y_j-x)^3}w_j}
    {\sum\limits_{j=1}^N \frac{(y_j-a)^{-1}}{y_j-x}w_j}, 
    \quad x \in \Omega^c.
\end{align}
The quadrature rule for $v''(x)$ requires $v(x)$ and $v'(x)$, and theses
are computed using the quadrature rules in
equations~\eqref{eqn:BaryvInterior},~\eqref{eqn:BaryvExterior},~\eqref{eqn:BaryvprimeInterior},
and~\eqref{eqn:BaryvprimeExterior}. In section~\ref{sec:results}, these
quadrature rules are used to form simulations of nearly-touching eroding
grains, and to study dynamics of the flow in regions arbitrarily close
to eroding grains.

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:intro}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Porous media flow plays an important role in many environmental and
industrial applications.  Depending on the application, length scales
can vary from $10^{-6}\mathrm{m}$ to
$10^{-1}\mathrm{m}$~\citep{mil-chr-imh-mcb-ped1998} and velocity scales
can be as small as $10^{-1}\mathrm{m/day}$~\citep{kut-scr-dav-ham1995}.
Moreover, for a single porous geometry, the pore sizes and velocities
can range over several orders of magnitude.  Numerical methods that
resolve this range of scales offer the ability to: (i) characterize
dispersion~\citep{saf1959}, (ii) quantify
mixing~\citep{leb-den-dav-bol-car-dec-bou2011, den-leb-eng-bij2011}, and
(iii) develop meaningful constitutive relationships that link the
microscopic and macroscopic realms~\citep{mil-chr-imh-mcb-ped1998}.
Examples of coarse-grained models for porous media flow include
permeability-porosity relationships~\citep{dar-mcc1998, car1937},
tortuosities~\citep{mat-kha-koz2008, dud-koz-mat2011, kop-kat-tim1996},
geometry connectivity~\citep{knu-car2005}, anomalous
dispersion~\citep{den-cor-sch-ber2004}, and more.

@@

Owing to the scales present in groundwater flow~\citep{bea1972}, we
model the hydrodynamics with the two-dimensional incompressible Stokes
equations.  Meanwhile, individual grains erode at a rate proportional to
the hydrodynamic shear stress~\citep{wan-fel2004,
ris-moo-chi-she-zha2012, moore2013self, par-izu2000}.  Since the fluid
equations are linear and homogeneous, they are converted to a boundary
integral equation (BIE). Therefore, we resolve the non-negligible
interactions between nearly-touching arbitrarily shaped bodies, and the
erosion rate is computed accurately. We also compute the vorticity in
the fluid bulk since, on solid boundaries, vorticity reduces to shear
and thus provides a convenient way to simultaneously visualize local
erosion rates and changes in the surrounding flow
(figure~\ref{fig:Eroding50vort}).

@@

To compute stable simulations of erosion, we use methods of high-order
in both space and time. The time integration is unchanged from previous
work~\citep{qua-moo2018}.  We apply a mild regularization and a
smoothing term to eliminate numerical instabilities that can be
triggered by changes in sign of the shear stress, and we use a stable
second-order Runge-Kutta method applied to the {\thL}
coordinates~\citep{hou-low-she1994} of the eroding grains.  In this
work, we introduce a new quadrature method to resolve dense suspensions.
The accuracy of the trapezoid rule, which was used in previous work, is
adequate for bodies that are sufficiently separated~\citep{tre-wei2014},
but not for grains in close contact.  One of the earliest quadrature
methods for nearly-singular integrands was developed
by~\citet{bak-she1986}, and in recent years, many other schemes have
followed~\citep{oja-tor2015, kli-tor2018, hel-oja2008a, bea-yin-wil2016,
bea-lai2001, klo-bar-gre-one2013}.  We use a Barycentric quadrature
method~\citep{bar2014, bar-wu-vee2015} since it is a non-intrusive
modification of the trapezoid rule, and the error is guaranteed to be
uniformly bounded.  We extend the original quadrature method to compute
the velocity gradient, which is needed to evaluate the shear stress and
the fluid vorticity.

To characterize transport through the resulting configurations, particle
trajectories must be computed. Depending on the application, microscale
transport can be modelled as pure advection~\citep{dea-qua-bir-jua2018,
leb-ded-dav-bou2007, cve-che-wen1996, puy-gou-den2019},
advection-diffusion~\citep{cus-hu-den1995, dag1987, den-ica-hid2018}, or
with a random walk~\citep{saf1959, bij-blu2006, ber-sch-sil2000}.  In
this work, we assume trajectories to be governed by pure advection
(ie.~no diffusion), so the particle trajectories are identical to the
streamlines. We compute trajectories $\ss(t)$ that are initialized at
$\ss_0$ by solving the advection equation
\begin{align}
  \frac{d\ss}{dt} = \uu(\ss,t), \quad \ss(0) = \ss_0,
  \label{eqn:tracers}
\end{align}
where $\uu$ is the fluid velocity.  Since there is no stiff diffusive
term, we solve~\eqref{eqn:tracers} with a fourth-order explicit
Runge-Kutta time stepping method. The Barycentric quadrature rule is
used to accurately compute trajectories that are close to an eroding
grain.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Governing Equations}
\label{sec:formulation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We start by defining the main variables used to model erosion.  We only
briefly summarize the model, and a more detailed description can be
found in previous work~\citep{qua-moo2018}.  We consider flows inside a
confined geometry $\Omega$ that contains $M$ eroding bodies with
boundaries $\gamma_\ell$, $\ell = 1,\ldots,M$.  The boundary of the
fluid domain is $\bd \Omega = \Gamma \cup \gamma_1 \cup \cdots \cup
\gamma_M$, where $\Gamma$ is the outer boundary, taken to be a slightly
smoothed version of the boundary of $[-3,3] \times [-1,1]$.  We only
place eroding bodies in $[-1,1] \times [-1,1]$ to create a buffer region
that allows the flow profile imposed at the inlet to transition to the more
complex flow intervening between the bodies. On each eroding body, $\nn$
is the normal vector pointing into the body, and $\ss$ is the unit
tangent vector pointing in the counterclockwise direction.  The normal
vector on $\Gamma$ is the outward unit normal.  Neglecting inertial
forces, the governing equations are
\begin{equation}
\label{eqn:erosionModel}
  \begin{split}
    \mu \Delta \uu = \grad p, &\hspace{20pt} \xx \in \Omega, \gap 
      &&\mbox{\em conservation of momentum}, \\
    \grad \cdot \uu = 0, &\hspace{20pt} \xx \in \Omega, \gap 
      &&\mbox{\em conservation of mass}, \\
    \uu = \mathbf{0}, &\hspace{20pt} \xx \in \gamma, \gap 
      &&\mbox{\em no slip on the eroding bodies}, \\
    \uu = \UU, &\hspace{20pt} \xx \in \Gamma, \gap 
      &&\mbox{\em outer wall velocity}, \\
    \Vn = \CE \, \abs{\tau}, &\hspace{20pt} \xx \in \gamma,
      &&\mbox{\em erosion model}.
  \end{split}
\end{equation}
Here, $\uu$ is the fluid velocity, $p$ is the pressure, $\UU$ is a
prescribed Hagen-Poiseuille velocity field, $\Vn$ is the normal velocity
of $\gamma$, and
\begin{align}
  \tau = -(\nabla \uu + \nabla \uu^T) \nn \cdot \ss
  \label{eqn:shearStress}
\end{align}
is the shear stress on $\gamma$. We simulate erosion by alternating
between solving the fluid equations and advancing the eroding grains.
The strength of $\UU$ is adjusted at each time step to achieve a
constant pressure drop across the channel, motivated by the geological
situation of a porous medium connecting two regions of  fixed hydraulic
heads.

@@

Once the velocity is computed in $\Omega$, the tortuosity can be
computed with the Eulerian velocity field~\eqref{eqn:tortuosity2}.  We
compute the velocity at $\xx_{ij} = (-1 + i\Delta x, -1 + j\Delta y)$,
$i,j=1,\ldots,N$, where $\Delta x = \Delta y = 2/N$, and the velocity at
points inside an eroding body are assigned a value of 0.  Then, the
tortuosity is approximated as
\begin{align}
  T = \left(\sum_{i,j=1}^N \|\uu(\xx_{ij})\|\Delta x \Delta y \right) 
      \Bigg/ 
      \left(\sum_{i,j=1}^N u_1(\xx_{ij}) \Delta x \Delta y \right)
%  T = \frac{\sum\limits_{i,j=1}^N \|\uu(\xx_{ij})\|\Delta x \Delta y}
%      {\sum\limits_{i,j=1}^N u_1(\xx_{ij}) \Delta x \Delta y}.
\end{align}

@@

Again, given a complex-valued density function $\eta$, the boundary data
of $v[\eta](x)$, as defined in equation~\eqref{eqn:laplaceComplex},
satisfies the Sokhotski-Plemelj jump relation~\eqref{eqn:SPrelation}.
Since the limiting boundary data of $v$ differs when considering $x \in
\Omega$ and $x \in \Omega^c$, we denote the boundary data as $v^-$ for
$x \in \Omega$, and as $v^+$ for $x \in \Omega^c$.  Rather than directly
applying the trapezoid rule to approximate $v(x)$ in
equation~\eqref{eqn:cauchy}, we start with the identity
\begin{align}
  \int_{\bd\Omega} \frac{v^{-}(y) - v(x)}{y - x} \,dy = 0,
    \quad x \in \Omega.
\end{align}
Since the integrand is bounded for all $x \in \Omega$, we can apply the
trapezoid rule
\begin{align}
  \sum_{j=1}^{N} \frac{v^{-}(y_j) - v(x)}{y_j - x} w_j \approx 0,
\end{align}
and the error is independent of $x$.  Rearranging for $v(x)$, we have
the interior Barycentric quadrature rule
\begin{align}
  v(x) = \left(\sum_{j=1}^N \frac{v^{-}(y_j)}{y_j - x} w_j \right)
  \Bigg/
  \left(\sum_{j=1}^N \frac{1}{y_j - x} w_j \right), 
  \quad x \in \Omega.
%  v(x) = \frac{\sum\limits_{j=1}^N \frac{v^{-}(y_j)}{y_j - x} w_j}
%  {\sum\limits_{j=1}^N \frac{1}{y_j - x} w_j}, \quad x \in \Omega.
  \label{eqn:BaryvInterior}
\end{align}
Using a similar construction and letting $a$ be any point inside
$\Omega$, the exterior Barycentric quadrature rule is
\begin{align}
  v(x) = \frac{1}{x-a} 
    \left(\sum_{j=1}^N \frac{v^+(y_j)}{y_j - x}w_j \right)
    \Bigg/
    \left(\sum_{j=1}^N \frac{(y_j - a)^{-1}}{y_j - x}w_j \right),
    \quad x \in \Omega^c.
%  v(x) = \frac{1}{x-a} 
%    \frac{\sum\limits_{j=1}^N \frac{v^+(y_j)}{y_j - x}w_j}
%    {\sum\limits_{j=1}^N \frac{(y_j - a)^{-1}}{y_j - x}w_j},
%    \quad x \in \Omega^c.
  \label{eqn:BaryvExterior}
\end{align}

Similar constructions can be used to derive quadrature rules
for $v'(x)$.  For $x \in \Omega$, we have the identity
\begin{align}
  \int_{\bd\Omega} \frac{v^-(y) - v(x) + (y-x)v'(x)}{(y - x)^2} = 0,
\end{align}
and the integrand is bounded for all $x \in \Omega$.  Therefore, after
applying the trapezoid rule and rearranging for $v'(x)$, we have the
interior Barycentric quadrature rule
\begin{align}
  v'(x) = \left(\sum_{j=1}^{N}
    \frac{v^{-}(y_j) - v(x)}{(y_j-x)^2} w_j \right)
  \Bigg/
  \left(\sum_{j=1}^{N} \frac{1}{y_j-x} w_j\right), 
  \quad x \in \Omega.
%  v'(x) = \frac{\sum\limits_{j=1}^{N}
%    \frac{v^{-}(y_j) - v(x)}{(y_j-x)^2} w_j}
%  {\sum\limits_{j=1}^{N} \frac{1}{y_j-x} w_j}, 
%  \quad x \in \Omega.
  \label{eqn:BaryvprimeInterior}
\end{align}
Using a similar construction, the exterior Barycentric quadrature rule
for the the first derivative is
\begin{align}
  v'(x) = \frac{1}{x-a} \left(\sum_{j=1}^N
    \frac{v^+_j - v(x)}{(y_j - x)^2} w_j \right)
    \Bigg/
    \left(\sum_{j=1}^N \frac{(y_j-a)^{-1}}{y_j - x} w_j\right),
    \quad x \in \Omega^c.
%  v'(x) = \frac{1}{x-a} \frac{\sum\limits_{j=1}^N
%    \frac{v^+_j - v(x)}{(y_j - x)^2} w_j}
%    {\sum\limits_{j=1}^N \frac{(y_j-a)^{-1}}{y_j - x} w_j},
%    \quad x \in \Omega^c.
  \label{eqn:BaryvprimeExterior}
\end{align}
Note that $v(x)$ is required to compute $v'(x)$ for both the interior
and exterior case, and this is available using the Barycentric
quadrature rules~\eqref{eqn:BaryvInterior}
and~\eqref{eqn:BaryvExterior}.

To compute the shear stress and vorticity, we require a Barycentric
quadrature rule for $v''(x)$.  The derivation is largely based on the
work of~\citet[see equation (2.12)]{ioa-pap-per1991}.  We start with
the second derivative of the Cauchy integral theorem
\begin{align}
  0 = \frac{1}{2\upi i} \int_{\bd\Omega} 
      \frac{2v^{-}(y)}{(y-x)^3}\,dy - v''(x).
\end{align}
For the interior case, $x \in \Omega$, we use the identity
\begin{align}
  \frac{1}{2\upi i}\int_{\bd\Omega} \frac{1}{(y-x)^n}\, dy = 
  \left\{
    \begin{array}{ll}
      1, & n = 1, \\
      0, & n = 2,3,\ldots. \\
    \end{array}
  \right.
\end{align}
Combining this identity with the Cauchy integral representation of
$v''(x)$, we have
\begin{align}
  0 = \frac{1}{2\upi i} \int_{\bd\Omega} 
      \frac{2v^{-}(y) - v''(x)(y-x)^2 - 2v(x) - 2(y-x)v'(x)}
      {(y-x)^3}\, dy.
\end{align}
This integrand is constructed so that it is bounded for all $x \in
\Omega$, and applying the trapezoid rule, we have
\begin{align}
  0 \approx  \sum_{j=1}^{N} 
      \frac{2v^{-}(y_j) - v''(x)(y_j-x)^2 - 2v(x) - 2(y_j-x)v'(x)}
      {(y_j-x)^3} w_j,
\end{align}
where the accuracy is independent of $x$.  Solving for $v''(x)$, the
Barycentric quadrature rule for the interior second derivative at $x \in
\Omega$ is
\begin{align}
  v''(x) \approx \left(2\sum_{j=1}^N 
    \frac{v^{-}_{j} - v(x) - (y_j-x)v'(x)}{(y_j-x)^3}w_j \right)
    \Bigg/
    \left(\sum_{j=1}^N \frac{1}{y_j-x}w_j\right).
%  v''(x) &\approx \frac{2\sum\limits_{j=1}^N 
%    \frac{v^{-}_{j} - v(x) - (y_j-x)v'(x)}{(y_j-x)^3}w_j}
%    {\sum\limits_{j=1}^N \frac{1}{y_j-x}w_j}, \quad x \in \Omega.
\end{align}
For the exterior case, $x \in \Omega^c$, we start with the identity
\begin{align}
\frac{1}{x-a} &= -\frac{1}{2\upi i}\int_{\bd\Omega} 
    \frac{(y-a)^{-1}}{y-x}\, dy. 
\end{align}
Combining this identity with the Cauchy integral representation of
$v''(x)$, we have
\begin{align}
  0 = \frac{1}{2\upi i} \int_{\bd\Omega} 
    \frac{2v^+(y) - 2v(x) - 2(y-x)v'(x) + (y-a)^{-1} (x-a) (y-x)^2 v''(x)}
    {(y-x)^3}.
\end{align}
As in the interior case, the integrand is chosen so that it is bounded
for all $x \in \Omega^c$.  Therefore, after applying the trapezoid rule
and solving for $v''(x)$, we have the Barycentric quadrature rule for
the exterior second derivative at $x \in \Omega^c$
\begin{align}
  v''(x) \approx \frac{1}{x-a}\left(2\sum\limits_{j=1}^N
    \frac{v^{+}_{j} - v(x) - (y_j-x)v'(x)}{(y_j-x)^3}w_j\right)
    \Bigg/
    \left(\sum_{j=1}^N \frac{(y_j-a)^{-1}}{y_j-x}w_j\right).
%  v''(x) &\approx \frac{1}{x-a}\frac{2\sum\limits_{j=1}^N
%    \frac{v^{+}_{j} - v(x) - (y_j-x)v'(x)}{(y_j-x)^3}w_j}
%    {\sum\limits_{j=1}^N \frac{(y_j-a)^{-1}}{y_j-x}w_j}, 
%    \quad x \in \Omega^c.
\end{align}
The quadrature rule for $v''(x)$ requires $v(x)$ and $v'(x)$, and theses
are computed using the quadrature rules in
equations~\eqref{eqn:BaryvInterior},~\eqref{eqn:BaryvExterior},~\eqref{eqn:BaryvprimeInterior},
and~\eqref{eqn:BaryvprimeExterior}. In section~\ref{sec:results}, these
quadrature rules are used to form simulations of nearly-touching eroding
grains, and to study dynamics of the flow in regions arbitrarily close
to eroding grains.

@@

% OUTDATED STUFF
%Examples of coarse-grained models for porous media flow include
%pressure-saturation-permeability
%relationships~\citep{mil-chr-imh-mcb-ped1998}, permeability-porosity
%relationships~\citep{dar-mcc1998, car1937, won-kop-tom1984}, first
%passage time distributions~\citep{ber-sch-sil2000, hym-den-hag-kan2019,
%cve-che-wen1996}, tortuosities~\citep{hak-com-den2019, mat-kha-koz2008,
%dud-koz-mat2011, kop-kat-tim1996}, concentration
%distributions~\citep{ica-den2019, bel-sal-rin1992}, pore and flow
%distributions~\citep{ali-par-wei-bre2017}, geometry
%connectivity~\citep{knu-car2005, wes-blo-gra2001}, anomalous
%dispersion~\citep{dea-qua-bir-jua2018, den-cor-sch-ber2004,
%sie-ili-pri-riv-gua2019}, and more.

%A variety of numerical methods for the fluid solver are available
%including lattice-gas or lattice Boltzmann~\citep{kop-kat-tim1996,
%dar-mcc1998}, finite volumes~\citep{suo-liu-gan2019, den-ica-hid2018,
%sie-ili-pri-riv-gua2019}, finite
%differences~\citep{leb-ded-dav-bou2007, knu-car2005}, and cellular
%automata~\citep{rot1988}.
